On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-21 manuscripts/PNAS/review/point_by_point_response.txt

We next respond to the reviewer's question about the mathematical "tightness" for the three types of quantities in our results : (a) required coding error eps(delta_1, delta_2) sufficient to guarantee uniqueness in terms of desired dictionary delta_1 and sparse representation error delta_2, (b) the constants C_1 and C_2 involved in our sufficient linear bound on eps(delta_1, delta_2) in terms of the noise, and (c) the number of sufficient samples N guaranteeing uniqueness.  With respect to (a), even with fixed dictionary, the recovery of sparse representations up to an error that is linear in the noise is essentially optimal (ACD13) (up to constants). Thus, the fact that we have a sufficient error in coding for uniqueness being linear for both the desired dictionary and sparse vector recovery error is optimal.  Although optimal, our solution to Problem 1 involves constants C_1 and C_2, and it would be interesting and relevant for practical applications to determine the minimal such constants. In the case of sparsity k=1, we have given an example showing optimality of constants in this case. Finally, with respect to (c), we regard this as a fundamental open problem in the field: how many sparsely generated samples determine the linear coding model?  (One reason to be skeptical that a polynomial suffice is again (TP14, T15).)  However, there is some good news -- if k is fixed or if the sparse codes \bar x_i are known to come from a polynomial number of supports, then a polynomial number of samples N is enough to guarantee uniqueness. We have revised the manuscript to discuss this last fact (which, incidentally, does not follow from the results of (HS15) in the noiseless case).

It is true, however, that we have not determined the tightest conditions regarding the constants and sample size necessary for uniqueness in sparse linear coding. We nevertheless believe that these sufficient conditions yield strong enough conclusions to stand on their own; for instance, in synchronizing device parameters for a particular sparsely representable sensory ensemble.

### 2017-08-21 manuscripts/PNAS/review/point_by_point_response.txt

To be clear, the new perspective we take on the problem yields the following powerful conclusions beyond those of a straightforward extension of (HS15) to the noisy case:

1) An extension to the case where the number of dictionary elements is unknown: The results of (HS15) only apply to the case where the matrix B has the same number of columns as A. We forgo this assumption and show that B must have at least as many columns as A and contains (up to noise) a subset of the columns of A. The size of this subset depends on a simple relation between the number of columns of B and the regularity of the support-set hypergraph.

2) A significant reduction in sample complexity: To identify the n x m generating dictionary A, (HS15) require that data be sampled from every k-dimensional subspace spanned by the m columns of A (that is, {m \choose k} subspaces in total). We show that the data need only be sampled from m subspaces in total (e.g. those supported by consecutive intervals of length k in some cyclic order on [m]) and in some cases as few as 2\sqrt{m} (e.g. when m = k^2, take the supports to be the rows and columns of [m] arranged in a square grid). Moreover, if the size of the support set of reconstructing codes is known to be polynomial in m and k, then the pigeonholing argument in the proof of Thm. 1 requires only a polynomial number of samples distributed over a polynomial number of supports; thus, N is then polynomial in \bar m, k. This point was only hinted at in the Discussion section of our original submission, but we have included it in the updated manuscript to make clear the power of our approach over that taken in (HS15).

3) No spark condition requirement for the generating matrix A: One of the mathematically significant contributions made by (HS15) was to forgo the constraint that the recovery matrix B also satisfy the spark condition, in contrast to all previously known uniqueness results which (either explicitly or implicitly) made this assumption.  Our proof is powerful enough to show that, in fact, even the matrix A need not satisfy the spark condition to be identifiable! Rather, it need only be injective on the union of subspaces with supports that form sets in a hypergraph satisfying the singleton intersection property (SIP). (For example, consider the matrix A = [e_1,.., e_5, v] where v = e_1 + e_3 + e_5, and take H to be the set of all consecutive pairs of [m] arranged in cyclic order. Then A satisfies the assumptions of Thm. 1 for dictionary recovery without satisfying the spark condition.) We had omitted this point in our original submission of the manuscript to keep things simple, but have decided to include this fact in our resubmission. This also required us to redefine the restricted matrix lower bound L_k to be in terms of a hypergraph H (L_H in the revised manuscript), which is an interesting object for further study in its own right.

### 2017-08-21 manuscripts/PNAS/review/point_by_point_response.txt

3) No spark condition requirement for the generating matrix A: One of the mathematically significant contributions made by (HS15) was to forgo the constraint that the recovery matrix B also satisfy the spark condition, in contrast to all previously known uniqueness results which (either explicitly or implicitly) made this assumption.  Our proof is powerful enough to show that, in fact, even the matrix A need not satisfy the spark condition to be identifiable! Rather, it need only be injective on the union of subspaces with supports that form sets in a hypergraph satisfying the singleton intersection property (SIP). (For example, consider the matrix A = [e_1,.., e_5, v] where v = e_1 + e_3 + e_5, and take H to be the set of all consecutive pairs of [m] arranged in cyclic order. Then A satisfies the assumptions of Thm. 1 for dictionary recovery without satisfying the spark condition.) We had omitted this point in our original submission of the manuscript to keep things simple, but have decided to include this fact in our resubmission. This also required us to redefine the restricted matrix lower bound L_k to be in terms of a hypergraph H (L_H in the revised manuscript), which is an interesting object for further study in its own right.

We must also reiterate here that our solution to Prob. 2, that of most interest to practitioners of dictionary learning methods, is to our knowledge the first of its kind in both the noise-free and noisy domains. We recognize that this was not clearly communicated in the Discussion section of our submitted manuscript.

@@

Let A satisfy eqn [2], x_1,..,x_N be a set of k-sparse vectors, and E \subseteq \choose{[m],k}, and \bar{m} be given and satisfy the following property:

For every S \in E, there exists a subset of x_1,..,x_N of at least size (k-1)\choose{\bar{m},k} with support in S. Moreover, any k vectors in this subset are linearly independent.

Is this what you meant?
-----------------

RESPONSE: Yes, though actually for every S \in E (now \mathcal{H} in the revised manuscript) there are "more than" (k-1)\choose{\bar{m},k} vectors with support in S (not "at least"). Thank you for helping us to clarify our statements. We have updated the manuscript to disambiguate these conditions.

### 2017-08-21 manuscripts/PNAS/review/point_by_point_response.txt

Is this what you meant?
-----------------

RESPONSE: Yes, though actually for every S \in E (now \mathcal{H} in the revised manuscript) there are "more than" (k-1)\choose{\bar{m},k} vectors with support in S (not "at least"). Thank you for helping us to clarify our statements. We have updated the manuscript to disambiguate these conditions.

-----------------
-- Related point: Is there a condition on \ell in Theorem 1?
-----------------

RESPONSE: Yes, thanks for this question. The condition on \ell is perhaps best understood as a condition on \bar m not being too large (given \ell) in order to guarantee the recovery of at least one dictionary element. We have updated the manuscript to reflect this, and the condition now reads "provided \bar m < \ell m / (\ell - 1)" where previously it read "provided p = \bar m − \ell( \bar m − m ) is positive". (Note: in the revised manuscript, we have actually opted to use r instead of \ell for the regularity of the hypergraph.)

@@

The paper's writing and composition is clear (with an exception detailed below), concise, and relevant to the topic. The analytical work appears to be correct to the best of my understanding, and the paper's combined use of several results in the literature to extend an existing dictionary learning uniqueness condition to a new dictionary learning stability condition appears to be original.

Some questions that remain and may be worth discussing deal with the practical impact of these results.

When is it possible to determine (numerically) if enough data has been gathered to obtain an accurate estimate of the underlying sparsity dictionary?
-----------------

RESPONSE: A statement of this very general nature requires that we calculate the probability that all of the conditions of Thm. 1 are satisfied by random data. We have provided an "almost certain" guarantee of this flavor in Cor 2, wherein the sparse vector supports are constrained to form a hypergraph satisfying the SIP. It would be useful and interesting to calculate furthermore the probability that random supports form a hypergraph satisfying the SIP. We have opted not to include these calculations due in part to space constraints, but also because we believe that what sets our work apart from the vast majority of results in the field is their deterministic nature, e.g. they do not depend on any kind of assumption about the particular random distribution from which the sparse coefficients or dictionary entries are drawn.

### 2017-08-21 manuscripts/PNAS/review/point_by_point_response.txt

The paper's writing and composition is clear (with an exception detailed below), concise, and relevant to the topic. The analytical work appears to be correct to the best of my understanding, and the paper's combined use of several results in the literature to extend an existing dictionary learning uniqueness condition to a new dictionary learning stability condition appears to be original.

Some questions that remain and may be worth discussing deal with the practical impact of these results.

When is it possible to determine (numerically) if enough data has been gathered to obtain an accurate estimate of the underlying sparsity dictionary?
-----------------

RESPONSE: A statement of this very general nature requires that we calculate the probability that all of the conditions of Thm. 1 are satisfied by random data. We have provided an "almost certain" guarantee of this flavor in Cor 2, wherein the sparse vector supports are constrained to form a hypergraph satisfying the SIP. It would be useful and interesting to calculate furthermore the probability that random supports form a hypergraph satisfying the SIP. We have opted not to include these calculations due in part to space constraints, but also because we believe that what sets our work apart from the vast majority of results in the field is their deterministic nature, e.g. they do not depend on any kind of assumption about the particular random distribution from which the sparse supports, coefficients, or dictionary entries are drawn.

### 2017-08-22 manuscripts/PNAS/chazthm_pnas.tex
final edits

The constant $C_2 = C_2(\mathbf{A}, \mathcal{H})$ is given in turn below, in terms of one used in \cite{Deutsch12} to analyze the convergence of the alternating projections algorithm for projecting a point onto an intersection of subspaces. 
We use it to bound the distance between a point and the intersection of subspaces given an upper bound on its distance from each individual subspace.

>>>>

The constant $C_2 = C_2(\mathbf{A}, \mathcal{H})$, in turn, will be presented relative to a quantity from \cite{Deutsch12} used to analyze the convergence of the alternating projections algorithm. 
More specifically, the following definition is used to bound the distance between a point and the intersection of subspaces given an upper bound on its distance from each individual subspace in Lem.~\ref{DistanceToIntersectionLemma}, below.
